A Case Study in Preserving a High Energy Physics Application

### 2014-01-29 paper.tex
Add little bits about the software used.

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is to short-lived to be detected
itself, it has to be reconstructed from its decay products.  The
analysis discussed in this section focuses on the Higgs decaying to
two tau leptons, which are themselves reconstructed from particle
showers.  Within the analysis, events are identified that show a
signature of decay products compatible with both hadronic tau and top
decays.  Properties of such events are used to distinguish signal-like
(matching simulated Higgs decays) and background-like events and are
also used in further statistical analysis.

@@

As AOD data are too large to be iteratively processed repetitively in
an physics analysis workflow, it is normally reduced further in
structural complexity and content.  For the analysis under
investigation here, this is a two-step process.  First, the AOD data
are processed at the Notre Dame working group cluster to BEAN events,
containing only trivial data containers packed in vectors.  This step
is time and CPU intensive and contains data of 11.6$\,$TB to be
analyzed by the tau analysis.  The BEAN format and data is shared
within the analysis group looking at Higgs production in association
with top quarks, which is formed by groups from a few American and
European universities.

In the second step, which is the beginning of the actual tau analysis,
the data are reduced to variables relevant to the tau analysis, while
only events matching basic quality criteria are kept.  This results in
a dataset of 43.3$\,$GB.  Again, the Notre Dame CMS groups cluster
resources are used to perform this reduction and selection, with
highly customized software.

The final data analysis, investigated below, can be run as a single
process, and contains a stringent event selection to keep only high
quality candidate events for the underlying physical process (using
122$\,$MiB of space).  Quantities from the selected events can be
both plotted and used in multivariate analysis to determine the level
of expected signal in real data.

>>>>

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is too short-lived to be detected
itself, it has to be reconstructed from its decay products.  The
analysis discussed in this section focuses on the Higgs decaying to
two tau leptons, which are themselves reconstructed from particle
showers.  Within the analysis, events are identified that show a
signature of decay products compatible with both hadronic tau and top
decays.  Properties of such events are used to distinguish signal-like
(matching simulated Higgs decays) and background-like events and are
also used in further statistical analysis.

@@

As AOD data are too large to be iteratively processed repetitively in
an physics analysis workflow, it is normally reduced further in
structural complexity and content.  For the analysis under
investigation here, this is a two-step process.  First, the AOD data
are processed at the Notre Dame working group cluster to BEAN events,
containing only trivial data containers packed in vectors.  This step
is time and CPU intensive and its output contains data of 11.6$\,$TB to be
analyzed by the tau analysis.
It is performed by a small custom code framework,
which is built on top of the CMS software stack, CMSSW,
and uses packages provided by several other special interest groups within CMS.
While the CMSSW framework is installed locally,
the various packages used are checked out from CVS,
and the BEAN framework is stored in git.
This is scheduled to change,
as the CMSSW distribution model switches to a virtual filesystem mounted via FUSE,
and special interest groups move their code to git.
The BEAN format, production code, and
data are shared within the analysis group looking at Higgs production
in association with top quarks, which is formed by groups from a few
American and European universities,
consisting of up to a few dozen contributors.

In the second step, which is the beginning of the actual tau analysis,
the data are reduced to variables relevant to the tau analysis, while
only events matching basic quality criteria are kept.  This results in
a dataset of 43.3$\,$GB.  Again, the Notre Dame CMS groups cluster
resources are used to perform this reduction and selection,
running highly customized software,
built on CMSSW and the BEAN framework,
but with output code written and maintained by a few people only.
Again, the code is stored in a git repository.

The final data analysis, investigated below, can be run as a single
process, and contains a stringent event selection to keep only high
quality candidate events for the underlying physical process (using
122$\,$MiB of space).  Quantities from the selected events can be
both plotted and used in multivariate analysis to determine the level
of expected signal in real data.
This package is written using the CMSSW build framework,
but only utilizes code from ROOT,
a particle physics toolkit underlying CMSSW,
and a few external python dependencies for convenience.
The latter have to be manually fetched and installed,
while the analysis program is built by CMSSW after being checked out of git.
